---
title: "Retrotransposons analysis 5 years survival"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library, echo = FALSE}
library(tidyverse)
library(labelled)
# library(REMP)
# library(ComplexHeatmap)
library(gtsummary)
library(survival)
library(survminer)
theme_gtsummary_compact()
theme_set(theme_classic())
```

<style>
div.darkblue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">Retrotransposons analysis in AACES and NCOCS data</span>

</div>
<br>

```{r load}
load(paste0(here::here(), "/cleaned_07082022.rda"))

cluster_res_list <- read_rds(paste0(here::here(), "/clustering/MOVICS/movics_cluster_res_list_08032023.rds"))

L1_imputed <- read_rds(paste0(here::here(), "/L1_imputed.rds"))
Alu_imputed <- read_rds(paste0(here::here(), "/Alu_imputed.rds"))
ERV_imputed <- read_rds(paste0(here::here(), "/ERV_imputed.rds"))

BRCA_mutations <- read_csv(paste0(here::here(), "/ancestry_and_mutation_3_12_24.csv"))
patient_id <- read_csv(paste0(here::here(), "/Patient ids  for BRCA and TP53 data request.csv"))
failed_qc <- read_csv(paste0(here::here(), "/Patient failed qc brca tp53 mutation.csv"))

survival_time_Nov2023 <- read_csv(paste0(here::here(), "/survival_time_Nov2023_2024-06-19.csv")) %>% 
  mutate(suid = as.character(suid))
```


```{r cleaning}
coca_cluster_results <- cluster_res_list$COCA$clust.res %>%
  remove_rownames() %>% 
  mutate(patient_id = str_remove(samID, "X")) %>% 
  select(-samID, coca_RE_cluster = clust)

phenoclean <- phenoclean %>%
  select(-c(starts_with("chr"))) %>%
  remove_rownames() %>%
  mutate(suid = Sample_Name) %>%
  select(patient_id = Complete.Barcode, suid, everything(), -c(Sample_Name, roi_num : Basename))

phenoclean <- phenoclean %>%
  mutate(bmi_cat = case_when(
    BMI_recent < 25                                         ~ "<25",
    BMI_recent >= 25 &
      BMI_recent < 30                                       ~ "25-29",
    BMI_recent >= 30 &
      BMI_recent < 35                                       ~ "30-34",
    BMI_recent >= 35                                        ~ "≥35"
  ), bmi_cat =
    factor(bmi_cat, levels = c("<25", "25-29", "30-34", "≥35"))) %>%
  mutate(bmi_cat2 = case_when(
    BMI_recent < 30                                         ~ "<30",
    BMI_recent >= 30                                        ~ "≥30"
  ), bmi_cat2 =
    factor(bmi_cat2, levels = c("<30", "≥30"))) %>%
  mutate(stage_cat = case_when(
    stage == "Distant"              ~ "Late",
    stage == "Localized"            ~ "Early",
    stage == "Regional"             ~ "Early"
  ), stage_cat = factor(stage_cat, levels = c("Early", "Late"))
  ) %>% 
  mutate(smoking_status = case_when(
    str_detect(smokcurrent, "current|former") ~ "Ever",
    smokcurrent == "never smoker"             ~ "Never"
  )) %>% 
  # Rename older vital status but keep in data
  rename(vitalstatus_2022 = vitalstatus,
         timelastfu_2022 = timelastfu) %>% 
  # Create survival var
  left_join(., survival_time_Nov2023, by = "suid") %>% 
  mutate(vitalstatus = case_when(
    os_event == 0                            ~ "Alive",
    os_event == 1                            ~ "Deceased",
    TRUE                                     ~ NA_character_
  )) %>% 
  mutate(os_time_5year = case_when(
    os_time_months <= 5 * 12      ~ os_time_months,
    os_time_months > 5 * 12       ~ 60,
    TRUE                                                       ~ NA_real_
  )) %>% 
  mutate(os_event_5year = case_when(
    os_time_months <= 60        ~ os_event,
    os_time_months > 60         ~ 0,
    TRUE                        ~ NA_real_
  ))

```

```{r BRCA cleaning}
failed_qc <- paste0(failed_qc$suid, collapse = "|")
BRCA_mutations <- BRCA_mutations %>% 
  full_join(patient_id, ., by = "suid") %>% 
  mutate(across(c(where(is.character), -Complete.Barcode), ~ str_to_lower(.))) %>% 
  mutate(suid = as.character(suid)) %>% 
  mutate_at(c("germline_mutation_BRCA1", "germline_mutation_BRCA2", 
              "tumor_mutation_BRCA1", "tumor_mutation_BRCA2",
              "germline_mutation_TP53", "tumor_mutation_TP53"), 
            ~ case_when(
              . == 1                                         ~ "Yes",
              . == 2                                         ~ "No",
              TRUE                                           ~ as.character(.)
            )
  ) %>% 
  mutate_at(c("germline_mutation_BRCA1", "germline_mutation_BRCA2", 
              "tumor_mutation_BRCA1", "tumor_mutation_BRCA2",
              "germline_mutation_TP53", "tumor_mutation_TP53"), 
            ~ case_when(
              str_detect(suid, failed_qc)                    ~ NA_character_,
              TRUE                                           ~ .
            )
  ) %>% 
  mutate(BRCA1_carrier = case_when(
    tumor_mutation_BRCA1 == "Yes" |
      germline_mutation_BRCA1 == "Yes"                       ~ "Yes",
    tumor_mutation_BRCA1 == "No" &
      germline_mutation_BRCA1 == "No"                        ~ "No"
  )) %>% 
  mutate(BRCA2_carrier = case_when(
    tumor_mutation_BRCA2 == "Yes" |
      germline_mutation_BRCA2 == "Yes"                       ~ "Yes",
    tumor_mutation_BRCA2 == "No" &
      germline_mutation_BRCA2 == "No"                        ~ "No"
  )) %>% 
  mutate(pathogenic_germline_BRCA1 = case_when(
    germline_mutation_BRCA1 == "Yes" &
      str_detect(germline_clinvar_BRCA1, "pathogenic")                 ~ "Yes",
    germline_mutation_BRCA1 == "Yes" &
      !str_detect(germline_clinvar_BRCA1, "pathogenic")                 ~ "No",
    germline_mutation_BRCA1 == "No"                          ~ "No"
  )) %>% 
  mutate(pathogenic_germline_BRCA2 = case_when(
    germline_mutation_BRCA2 == "Yes" &
      str_detect(germline_clinvar_BRCA2, "pathogenic")                 ~ "Yes",
    germline_mutation_BRCA2 == "Yes" &
      !str_detect(germline_clinvar_BRCA2, "pathogenic")                 ~ "No",
    germline_mutation_BRCA2 == "No"                          ~ "No"
  )) %>% 
  mutate(pathogenic_tumor_BRCA1 = case_when(
    tumor_mutation_BRCA1 == "Yes" &
      str_detect(tumor_clinvar_BRCA1, "pathogenic")                    ~ "Yes",
    tumor_mutation_BRCA1 == "Yes" &
      !str_detect(tumor_clinvar_BRCA1, "pathogenic")                    ~ "No",
    tumor_mutation_BRCA1 == "No"                             ~ "No"
  )) %>% 
  mutate(pathogenic_tumor_BRCA2 = case_when(
    tumor_mutation_BRCA2 == "Yes" &
      str_detect(tumor_clinvar_BRCA2, "pathogenic")                    ~ "Yes",
    tumor_mutation_BRCA2 == "Yes" &
      !str_detect(tumor_clinvar_BRCA2, "pathogenic")                    ~ "No",
    tumor_mutation_BRCA2 == "No"                             ~ "No"
  )) %>% 
  mutate(pathogenic_germline_TP53 = case_when(
    germline_mutation_TP53 == "Yes" &
      str_detect(germline_clinvar_TP53, "pathogenic")        ~ "Yes",
    germline_mutation_TP53 == "Yes" &
      !str_detect(germline_clinvar_TP53, "pathogenic")       ~ "No",
    germline_mutation_TP53 == "No"                           ~ "No"
  )) %>% 
  mutate(pathogenic_tumor_TP53 = case_when(
    tumor_mutation_TP53 == "Yes" &
      str_detect(tumor_clinvar_TP53, "pathogenic")           ~ "Yes",
    tumor_mutation_TP53 == "Yes" &
      !str_detect(tumor_clinvar_TP53, "pathogenic")          ~ "No",
    tumor_mutation_TP53 == "No"                              ~ "No"
  )) %>%
  # mutate(germline_BRCA12 = case_when(
  #   germline_mutation_BRCA1 == "Yes" |
  #     germline_mutation_BRCA2 == "Yes"                       ~ "No",
  #   germline_mutation_BRCA1 == "No" &
  #     germline_mutation_BRCA2 == "No"                        ~ "No"
  # )) %>% 
  # mutate(tumor_BRCA12 = case_when(
  #   tumor_mutation_BRCA1 == "Yes" |
  #     tumor_mutation_BRCA2 == "Yes"                          ~ "No",
  #   tumor_mutation_BRCA1 == "No" &
  #     tumor_mutation_BRCA2 == "No"                           ~ "No"
  # )) %>% 
  mutate_at(c("germline_mutation_BRCA1", "germline_mutation_BRCA2", 
              "tumor_mutation_BRCA1", "tumor_mutation_BRCA2",
              "germline_mutation_TP53", "tumor_mutation_TP53",
              "pathogenic_germline_BRCA1", "pathogenic_germline_BRCA2",
              "pathogenic_tumor_BRCA1", "pathogenic_tumor_BRCA2",
              "pathogenic_germline_TP53", "pathogenic_tumor_TP53",
              "BRCA1_carrier", "BRCA2_carrier"),
            ~ factor(., levels = c("No",
                                   "Yes")))
```

```{r prep data}
met_data <- phenoclean %>% 
  inner_join(., coca_cluster_results,
            by = "patient_id") %>% 
  left_join(., BRCA_mutations, 
            by = c("patient_id" = "Complete.Barcode", "suid")) %>% 
  mutate(across(c(where(is.character), -patient_id), ~ str_to_sentence(.))) %>% 
  mutate(coca_RE_cluster = as.character(coca_RE_cluster))

write_rds(met_data, "met_data_01-23-2025.rds")
```

```{r Labeling}
var_label(met_data) <- list(refage = "Age at diagnosis",
                            stage_cat = "Stage",
                            debulking_CA125 = "Debulking Status", 
                            neoadjuvant = "Neoadjuvant chemotherapy",
                            adjuvant = "Adjuvant chemotherapy",
                            smokcurrent = "Current smoking status",
                            germline_mutation_BRCA1 = "BRCA1 germline carrier", 
                            germline_mutation_BRCA2 = "BRCA2 germline carrier", 
                            tumor_mutation_BRCA1 = "BRCA1 tumor carrier", 
                            tumor_mutation_BRCA2 = "BRCA2 tumor carrier",
                            germline_mutation_TP53 = "TP53 germline carrier", 
                            tumor_mutation_TP53 = "TP53 tumor carrier",
                            pathogenic_germline_BRCA1 = "pathogenic BRCA1 germline carrier", 
                            pathogenic_germline_BRCA2 = "pathogenic BRCA2 germline carrier",
                            pathogenic_tumor_BRCA1 = "pathogenic BRCA1 tumor carrier", 
                            pathogenic_tumor_BRCA2 = "pathogenic BRCA2 tumor carrier",
                            pathogenic_germline_TP53 = "pathogenic TP53 germline carrier", 
                            pathogenic_tumor_TP53 = "pathogenic TP53 tumor carrier",
                            coca_RE_cluster = "COCA cluster")
```

# I. Table 1. Patient characteristics overall
```{r}
met_data %>% 
  select(refage, race, ethnicity_dnam,
         karyotype,
         stage, stage_cat,
         histology, grade, 
         BMI_recent, bmi_cat, bmi_cat2,
         smokcurrent, smoking_status,
         debulking_CA125, neoadjuvant, adjuvant,
         vitalstatus2, os_event_5year,
         os_time_months, os_time_5year,
         coca_RE_cluster
  ) %>% 
  tbl_summary(
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    # label = list(refage ~ "Age at diagnosis",
    #              debulking_CA125 ~ "Debulking Status", 
    #              neoadjuvant ~ "Neoadjuvant chemotherapy",
    #              adjuvant ~ "Adjuvant chemotherapy",
    #              smokcurrent ~ "Current smoking status"
    # ),
    type = list(c(neoadjuvant, adjuvant) ~ "categorical"),
    digits = list(all_continuous() ~ 1)) %>% 
  bold_labels() %>% add_stat_label() %>%
  modify_header(list(label ~ "**Patient characteristics**", 
                     all_stat_cols() ~ "**{level}**, N = {n}"))

met_data %>% 
  select(contains("BRCA1"),
         contains("BRCA2"),
         contains("TP53")
  ) %>% 
  tbl_summary(
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    type = everything() ~ "categorical",
    digits = list(all_continuous() ~ 1)) %>% 
  bold_labels() %>% add_stat_label() %>%
  modify_header(list(label ~ "**Mutations characteristics**", 
                     all_stat_cols() ~ "**{level}**, N = {n}"))
```
<br>
<br>

# II. Table 2. Patient characteristics by cluster
```{r}
met_data %>% 
  select(refage, race, ethnicity_dnam,
         karyotype,
         stage, stage_cat,
         histology, grade, 
         BMI_recent, bmi_cat, bmi_cat2,
         smokcurrent, smoking_status,
         debulking_CA125, neoadjuvant, adjuvant,
         vitalstatus2, os_event_5year,
         coca_RE_cluster
  ) %>% 
  tbl_summary(by = coca_RE_cluster,
              statistic = list(all_continuous() ~ "{mean} ({sd})"),
              # label = list(refage ~ "Age at diagnosis",
              #              debulking_CA125 ~ "Debulking Status", 
              #              neoadjuvant ~ "Neoadjuvant chemotherapy",
              #              adjuvant ~ "Adjuvant chemotherapy",
              #              smokcurrent ~ "Current smoking status"
              # ),
              type = list(c(neoadjuvant, adjuvant) ~ "categorical"),
              digits = list(all_continuous() ~ 1)) %>%
  bold_labels() %>% add_stat_label() %>% 
  add_p() %>% bold_p(t=.05) %>% add_overall() %>% 
  modify_spanning_header(c(stat_1, stat_2) ~ "COCA cluster") %>% 
  modify_header(list(label ~ "**Patient characteristics**"))

met_data %>% 
  select(contains("BRCA1"),
         contains("BRCA2"),
         contains("TP53"),
         coca_RE_cluster
  ) %>% 
  tbl_summary(by = coca_RE_cluster,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    type = everything() ~ "categorical",
    digits = list(all_continuous() ~ 1)) %>% 
  bold_labels() %>% add_stat_label() %>%
  add_p() %>% bold_p(t=.05) %>% add_overall() %>% 
  modify_spanning_header(c(stat_1, stat_2) ~ "COCA cluster") %>% 
  modify_header(list(label ~ "**Mutations characteristics**", 
                     all_stat_cols() ~ "**{level}**, N = {n}"))
```

# III. Adjusted hazard ratios and 95% confidence intervals

## 1. Adjusted hazard ratios and 95% confidence intervals
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
PH assumption
```{r}
coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat,
      data = met_data)  %>%
  cox.zph()
```
<br>
<br>


## 2. HR adjusted by debulking status
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, debulking_CA125) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + debulking_CA125,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
PH assumption
```{r}
coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + debulking_CA125,
      data = met_data)  %>%
  cox.zph()
```
<br>
<br>

## 3. HR adjusted by BRCA1
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, germline_mutation_BRCA1) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + germline_mutation_BRCA1,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, pathogenic_germline_BRCA1) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + pathogenic_germline_BRCA1,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, tumor_mutation_BRCA1) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + tumor_mutation_BRCA1,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, pathogenic_tumor_BRCA1) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + pathogenic_tumor_BRCA1,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

## 4. HR adjusted by BRCA2
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, germline_mutation_BRCA2) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + germline_mutation_BRCA2,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, pathogenic_germline_BRCA2) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + pathogenic_germline_BRCA2,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, tumor_mutation_BRCA2) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + tumor_mutation_BRCA2,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, pathogenic_tumor_BRCA2) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + pathogenic_tumor_BRCA2,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

## HR adjusted by BRCA1 + BRCA2
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, germline_mutation_BRCA1, 
         germline_mutation_BRCA2) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + 
                germline_mutation_BRCA1 + germline_mutation_BRCA2,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, tumor_mutation_BRCA1, 
         tumor_mutation_BRCA2) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + 
                tumor_mutation_BRCA1 + tumor_mutation_BRCA2,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

Adjusted by BRCA1 overall + BRCA2 overall
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, BRCA1_carrier, 
         BRCA2_carrier) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + 
                BRCA1_carrier + BRCA2_carrier,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

## 5. HR adjusted by TP53
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, germline_mutation_TP53) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + germline_mutation_TP53,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, pathogenic_germline_TP53) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + pathogenic_germline_TP53,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, tumor_mutation_TP53) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + tumor_mutation_TP53,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```
```{r}
tbl1 <- met_data %>%
  select(coca_RE_cluster,
         refage, stage_cat, pathogenic_tumor_TP53) %>%
  tbl_uvregression(method = survival::coxph,
                   y = (Surv(time = met_data$os_time_5year,
                             event = met_data$os_event_5year)),
                   exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level") %>%
  bold_labels() %>% italicize_levels()
tbl2 <- coxph(Surv(time = met_data$os_time_5year,
                   event = met_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + pathogenic_tumor_TP53,
              data = met_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

tbl_merge(list(tbl1, tbl2), tab_spanner = c("**Univariable**", "**Multivariable**"))
```

# Forest plot

# IV. KM
```{r Survival, fig.height = 7}
ggsurvplot(survfit(Surv(os_time_5year, os_event_5year) ~ coca_RE_cluster,
                   data=met_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) #+ guides(colour = guide_legend(ncol = 1))
```
<br>
<br>

# V. Sensitivity analysis

<!-- ## 1. Sensitivity analysis in patients who received adjuvant -->
<!-- ```{r} -->
<!-- adj_data <- met_data %>% -->
<!--   filter(adjuvant == "Yes") -->

<!-- coxph(Surv(time = adj_data$os_time_5year, -->
<!--                    event = adj_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + debulking_CA125, -->
<!--               data = adj_data)  %>% -->
<!--   tbl_regression(exponentiate = TRUE) %>% -->
<!--   bold_p(t = .05) %>% -->
<!--   add_nevent(location = "level") %>% add_n(location = "level") -->
<!-- ``` -->
<!-- <br> -->
<!-- <br> -->

## 1. Sensitivity analysis in patients with distant (late) stage
```{r}
distant_data <- met_data %>%
  filter(stage_cat == "Late")

coxph(Surv(time = distant_data$os_time_5year,
                   event = distant_data$os_event_5year) ~ coca_RE_cluster + refage + debulking_CA125,
              data = distant_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_labels() %>% italicize_levels() %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")

ggsurvplot(survfit(Surv(os_time_5year, os_event_5year) ~ coca_RE_cluster,
                   data=distant_data),
           title = "OS Analysis",
           font.main = c(20, "bold", "black"),
           font.x = c(18, "bold", "black"),
           font.y = c(18, "bold", "black"),
           font.legend = c(16, "black"),
           font.tickslab = c(16, "bold", "black"),
           size = 1.5,

           xlab = "Time (months)",
           legend = "top",
           legend.title = "",
           pval = TRUE,
           conf.int = FALSE,
           # Censor
           censor = TRUE
) #+ guides(colour = guide_legend(ncol = 1))
```

## 2. Sensitivity analysis in patients with serous histology
```{r}
serous_data <- met_data %>%
  filter(histology == "Serous")

coxph(Surv(time = serous_data$os_time_5year,
                   event = serous_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + debulking_CA125,
              data = serous_data)  %>%
  tbl_regression(exponentiate = TRUE) %>%
  bold_labels() %>% italicize_levels() %>%
  bold_p(t = .05) %>%
  add_nevent(location = "level") %>% add_n(location = "level")
```

<!-- ## 3. Sensitivity analysis in patients with Black or african american DNA -->
<!-- ```{r} -->
<!-- black_data <- met_data %>% -->
<!--   filter(ethnicity_dnam == "Black or african american") -->

<!-- coxph(Surv(time = black_data$os_time_5year, -->
<!--                    event = black_data$os_event_5year) ~ coca_RE_cluster + refage + stage_cat + debulking_CA125, -->
<!--               data = black_data)  %>% -->
<!--   tbl_regression(exponentiate = TRUE) %>% -->
<!--   bold_p(t = .05) %>% -->
<!--   add_nevent(location = "level") %>% add_n(location = "level") -->
<!-- ``` -->





