---
title: "RE methylation cluster and immune data"
author: "Christelle Colin-Leitzinger"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: false
    theme: united
    highlight: pygments
    df_print: paged
editor_options: 
  chunk_output_type: console
---

<style type="text/css">

.figure {
   margin-top: 25px;
   margin-bottom: 100px;
}

table {
    margin-top: 10px;
    margin-bottom: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```

```{r library, echo = FALSE}
library(tidyverse)
library(labelled)
# library(REMP)
# library(ComplexHeatmap)
library(gtsummary)
library(survival)
library(survminer)
theme_gtsummary_compact()
theme_set(theme_classic())
```

<style>
div.darkblue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">RE methylation cluster and HiTIMED data</span>

</div>
<br>

```{r load}
load(paste0(here::here(), "/cleaned_07082022.rda"))

cluster_res_list <- read_rds(paste0(here::here(), "/clustering/MOVICS/movics_cluster_res_list_08032023.rds"))

# L1_imputed <- read_rds(paste0(here::here(), "/L1_imputed.rds"))
# Alu_imputed <- read_rds(paste0(here::here(), "/Alu_imputed.rds"))
# ERV_imputed <- read_rds(paste0(here::here(), "/ERV_imputed.rds"))

met_data <- read_rds(paste0(here::here(), "/met_data_01-23-2025.rds"))

hit_data <- readxl::read_xlsx(paste0(here::here(), "/HiTIMED_round.xlsx"))
```

```{r join HIT}
met_data <- left_join(met_data, hit_data, by = c("patient_id" = "Complete.Barcode")) %>% 
  mutate(CD4 = CD4mem + CD4nv) %>% 
  mutate(CD8 = CD8mem + CD8nv)
```












```{r Labeling}
var_label(met_data) <- list(refage = "Age at diagnosis",
                            stage_cat = "Stage",
                            debulking_CA125 = "Debulking Status", 
                            neoadjuvant = "Neoadjuvant chemotherapy",
                            adjuvant = "Adjuvant chemotherapy",
                            smokcurrent = "Current smoking status",
                            germline_mutation_BRCA1 = "BRCA1 germline carrier", 
                            germline_mutation_BRCA2 = "BRCA2 germline carrier", 
                            tumor_mutation_BRCA1 = "BRCA1 tumor carrier", 
                            tumor_mutation_BRCA2 = "BRCA2 tumor carrier",
                            germline_mutation_TP53 = "TP53 germline carrier", 
                            tumor_mutation_TP53 = "TP53 tumor carrier",
                            pathogenic_germline_BRCA1 = "pathogenic BRCA1 germline carrier", 
                            pathogenic_germline_BRCA2 = "pathogenic BRCA2 germline carrier",
                            pathogenic_tumor_BRCA1 = "pathogenic BRCA1 tumor carrier", 
                            pathogenic_tumor_BRCA2 = "pathogenic BRCA2 tumor carrier",
                            pathogenic_germline_TP53 = "pathogenic TP53 germline carrier", 
                            pathogenic_tumor_TP53 = "pathogenic TP53 tumor carrier",
                            coca_RE_cluster = "COCA cluster")
```

# I. Table 
```{r}
met_data %>% 
  select(Tumor : CD8nv, CD4, CD8,
         coca_RE_cluster
  ) %>% 
  tbl_summary(by = coca_RE_cluster,
              statistic = list(all_continuous() ~ "{mean} ({sd})"),
              type = list(c(CD4nv) ~ "continuous")#,
              # digits = list(all_continuous() ~ 1)
              ) %>%
  bold_labels() %>% add_stat_label() %>% 
  add_p() %>% bold_p(t=.05) %>% add_overall() %>% 
  modify_spanning_header(c(stat_1, stat_2) ~ "COCA cluster") %>% 
  modify_header(list(label ~ "**Patient characteristics**"))
```

## Plot
```{r}
met_data %>% 
  ggplot(aes(x = CD8mem, fill = coca_RE_cluster))+
  geom_histogram()

met_data %>% 
  ggplot(aes(x = CD8nv, fill = coca_RE_cluster))+
  geom_histogram()

met_data %>% 
  ggplot(aes(x = CD8mem, fill = coca_RE_cluster))+
  geom_density(alpha = 0.5)

met_data %>% 
  ggplot(aes(x = CD8nv, fill = coca_RE_cluster))+
  geom_density(alpha = 0.5)

met_data %>% 
  ggplot(aes(x = CD4, fill = coca_RE_cluster))+
  geom_density(alpha = 0.5)
```

# II. Suplemental figure?
### If only show T cell
```{r}
met_data %>% 
  ggplot(aes(x = CD8, fill = coca_RE_cluster))+
  geom_density(alpha = 0.5)
```

### If we want to show more cells
<!-- this is with free y axis -->
<!-- ```{r} -->
<!-- met_data %>%  -->
<!--   select(Monocyte : Bmem, Treg, -->
<!--          CD4, CD8, -->
<!--          coca_RE_cluster -->
<!--   ) %>%  -->
<!--   pivot_longer(cols = -coca_RE_cluster,  -->
<!--                names_to = "cell_type",  -->
<!--                values_to = "cell_count") %>%  -->
<!--   mutate(all = "All") %>%  -->
<!--   filter(!is.na(cell_count)) %>%  -->
<!--   ggplot(aes(x = all, y = cell_count, color = coca_RE_cluster))+ -->
<!--   geom_boxplot()+ -->
<!--   labs(x = "", y = "Cell Count")+ -->
<!--   scale_color_manual(values = c("red", "blue"), -->
<!--                      aesthetics = c("colour", "fill"), -->
<!--                      labels = c("Active", "Repressed"),  -->
<!--                      name = "COCA cluster")+ -->
<!--   facet_wrap(cell_type ~ ., scales = "free_y")+ -->
<!--   theme(axis.text.x = element_blank(), -->
<!--         axis.ticks.x = element_blank()) -->

<!-- met_data %>%  -->
<!--   select(Monocyte : Bmem, Treg, -->
<!--          CD4, CD8, -->
<!--          coca_RE_cluster -->
<!--   ) %>%  -->
<!--   pivot_longer(cols = -coca_RE_cluster,  -->
<!--                names_to = "cell_type",  -->
<!--                values_to = "cell_count") %>%  -->
<!--   mutate(all = "All") %>%  -->
<!--   filter(!is.na(cell_count)) %>%  -->
<!--   ggplot(aes(x = all, y = cell_count, color = coca_RE_cluster))+ -->
<!--   ggtitle("using the 'width' option for violin plot = all violins have the same maximum width")+ -->
<!--   geom_violin(scale = "width")+ -->
<!--   labs(x = "", y = "Cell Count")+ -->
<!--   scale_color_manual(values = c("red", "blue"), -->
<!--                      aesthetics = c("colour", "fill"), -->
<!--                      labels = c("Active", "Repressed"),  -->
<!--                      name = "COCA cluster")+ -->
<!--   facet_wrap(cell_type ~ ., scales = "free_y")+ -->
<!--   theme(axis.text.x = element_blank(), -->
<!--         axis.ticks.x = element_blank()) -->

<!-- met_data %>%  -->
<!--   select(Monocyte : Bmem, Treg, -->
<!--          CD4, CD8, -->
<!--          coca_RE_cluster -->
<!--   ) %>%  -->
<!--   pivot_longer(cols = -coca_RE_cluster,  -->
<!--                names_to = "cell_type",  -->
<!--                values_to = "cell_count") %>%  -->
<!--   mutate(all = "All") %>%  -->
<!--   filter(!is.na(cell_count)) %>%  -->
<!--   ggplot(aes(x = all, y = cell_count, color = coca_RE_cluster))+ -->
<!--   ggtitle("using the 'count' option for violin plot = areas are scaled proportionally to the number of observations")+ -->
<!--   geom_violin(scale = "count")+ -->
<!--   labs(x = "", y = "Cell Count")+ -->
<!--   scale_color_manual(values = c("red", "blue"), -->
<!--                      aesthetics = c("colour", "fill"), -->
<!--                      labels = c("Active", "Repressed"),  -->
<!--                      name = "COCA cluster")+ -->
<!--   facet_wrap(cell_type ~ ., scales = "free_y")+ -->
<!--   theme(axis.text.x = element_blank(), -->
<!--         axis.ticks.x = element_blank()) -->

<!-- met_data %>%  -->
<!--   select(#Monocyte : Bmem,  -->
<!--          Treg, CD4, CD8, -->
<!--          coca_RE_cluster -->
<!--   ) %>%  -->
<!--   pivot_longer(cols = -coca_RE_cluster,  -->
<!--                names_to = "cell_type",  -->
<!--                values_to = "cell_count") %>%  -->
<!--   mutate(all = "All") %>%  -->
<!--   filter(!is.na(cell_count)) %>%  -->
<!--   ggplot(aes(x = all, y = cell_count, color = coca_RE_cluster))+ -->
<!--   ggtitle("violin + boxplot")+ -->
<!--   geom_violin()+ -->
<!--   geom_boxplot(width=0.1, position = position_dodge(0.9), alpha=0.2, color = "grey", aes(group = interaction(coca_RE_cluster, cell_type)))+ -->
<!--   labs(x = "", y = "Cell Count")+ -->
<!--   scale_color_manual(values = c("red", "blue"), -->
<!--                      aesthetics = c("colour", "fill"), -->
<!--                      labels = c("Active", "Repressed"),  -->
<!--                      name = "COCA cluster")+ -->
<!--   facet_wrap(cell_type ~ ., ncol = 1, -->
<!--              scales = "free_x", switch = "y", axes = "all_x")+ -->
<!--   coord_flip() -->
<!--   theme(axis.text.y = element_blank(), -->
<!--         axis.ticks.y = element_blank()) -->
<!-- ``` -->






```{r}
met_data %>% 
  select(
         Treg, CD4, CD8,
         coca_RE_cluster
  ) %>% 
  pivot_longer(cols = -coca_RE_cluster, 
               names_to = "cell_type", 
               values_to = "cell_count") %>% 
  # mutate(all = "All") %>% 
  filter(!is.na(cell_count)) %>% 
  ggplot(aes(x = cell_type, y = cell_count, color = coca_RE_cluster))+
  ggtitle("long violin + boxplot")+
  geom_violin(scale = "width")+
  # ylim(0, 1)
  # geom_boxplot(width=0.1, position = position_dodge(0.9), alpha=0.2, color = "grey", aes(group = interaction(coca_RE_cluster, cell_type)))+
  # labs(x = "", y = "Cell Count")+
  scale_color_manual(values = c("red", "blue"),
                     aesthetics = c("colour", "fill"),
                     labels = c("Active", "Repressed"),
                     name = "COCA cluster")
  # coord_flip()+
  facet_wrap(cell_type ~ .)
  
met_data %>% 
  select(#Monocyte : Bmem, 
         Treg, CD4, CD8,
         coca_RE_cluster
  ) %>% 
  pivot_longer(cols = -coca_RE_cluster, 
               names_to = "cell_type", 
               values_to = "cell_count") %>% 
  # mutate(all = "All") %>% 
  filter(!is.na(cell_count)) %>% 
  ggplot(aes(x = cell_type, y = cell_count, color = coca_RE_cluster))+
  ggtitle("long violin + boxplot")+
  geom_violin(scale = "width")+
  # ylim(0, 1)
  # geom_boxplot(width=0.1, position = position_dodge(0.9), alpha=0.2, color = "grey", aes(group = interaction(coca_RE_cluster, cell_type)))+
  # labs(x = "", y = "Cell Count")+
  scale_color_manual(values = c("red", "blue"),
                     aesthetics = c("colour", "fill"),
                     labels = c("Active", "Repressed"),
                     name = "COCA cluster")
  # coord_flip()+
  facet_wrap(cell_type ~ .)
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
  
  met_data %>% 
  select(#Monocyte : Bmem, 
         Treg, CD4, CD8,
         coca_RE_cluster
  ) %>% 
  pivot_longer(cols = -coca_RE_cluster, 
               names_to = "cell_type", 
               values_to = "cell_count") %>% 
  mutate(all = "All") %>%
  filter(!is.na(cell_count)) %>% 
  ggplot(aes(x = all, y = cell_count, color = coca_RE_cluster))+
  ggtitle("violin + boxplot")+
  geom_violin(scale = "width")+
  # ylim(0, 1)
  geom_boxplot(width=0.1, position = position_dodge(0.9), alpha=0.2, color = "grey", aes(group = interaction(coca_RE_cluster, cell_type)))+
  # labs(x = "", y = "Cell Count")+
  scale_color_manual(values = c("red", "blue"),
                     aesthetics = c("colour", "fill"),
                     labels = c("Active", "Repressed"),
                     name = "COCA cluster")+
  # coord_flip()+
  facet_wrap(cell_type ~ .)
  
  
  met_data %>% 
  select(Treg, CD4T = CD4, CD8T = CD8,
         coca_RE_cluster
  ) %>% 
  pivot_longer(cols = -coca_RE_cluster, 
               names_to = "cell_type", 
               values_to = "cell_count") %>% 
  mutate(all = "All") %>%
  filter(!is.na(cell_count)) %>% 
  ggplot(aes(x = cell_type, y = cell_count, color = coca_RE_cluster))+
  geom_violin(scale = "width")+
  # ylim(0, 17)+
  geom_boxplot(width=0.1, position = position_dodge(0.9), alpha=0.2, color = "darkgrey", 
               aes(group = interaction(coca_RE_cluster, cell_type)))+
  labs(x = "", y = "Cell proportion estimate")+
    ylim(0, 20)+
  scale_color_manual(values = c("red", "blue"),
                     aesthetics = c("colour", "fill"),
                     labels = c("Active", "Repressed"),
                     name = "COCA cluster") + 
    stat_compare_means(label = "p.format", show.legend = FALSE, size = 3, label.y = 18)
  
  ggsave("Figure S6 immune cell proportion.pdf",
       width = 5,
       height = 5, 
       dpi = 600)
  
  theme_set(theme_grey()) # Call theme_grey then theme_classic in the plot code to eliminate ggbreak extra outer lines
  
  library(ggbreak)
  met_data %>% 
  select(Treg, CD4T = CD4, CD8T = CD8,
         coca_RE_cluster
  ) %>% 
  pivot_longer(cols = -coca_RE_cluster, 
               names_to = "cell_type", 
               values_to = "cell_count") %>% 
  mutate(all = "All") %>%
  filter(!is.na(cell_count)) %>% 
  ggplot(aes(x = cell_type, y = cell_count, color = coca_RE_cluster))+
  geom_violin(scale = "width")+
  # ylim(0, 17)+
  geom_boxplot(width=0.1, position = position_dodge(0.9), alpha=0.2, color = "darkgrey", 
               aes(group = interaction(coca_RE_cluster, cell_type)))+
  labs(x = "", y = "Cell proportion estimate", tag = "Figure S6")+
  scale_color_manual(values = c("red", "blue"),
                     aesthetics = c("colour", "fill"),
                     labels = c("Active", "Repressed"),
                     name = "COCA cluster")+
    expand_limits(y = c(0, 18))+
    # scale_y_cut(breaks = c(10.5))
    scale_y_break(c(10.5, 16), 
                  ticklabels=c(16, 18)
                  ) +
    stat_compare_means(label = "p.format", show.legend = FALSE, size = 3, label.y = 17.2)+
    theme_classic()+
    theme(plot.title.position = "plot")

ggsave("Figure S6 immune cell proportion y break with title.pdf",
       width = 6,
       height = 5, 
       dpi = 600)

theme_set(theme_classic())
```






<!-- this is with fix axis -->
<!-- ```{r} -->
<!-- met_data %>%  -->
<!--   select(Monocyte : Bmem, Treg, -->
<!--          CD4, CD8, -->
<!--          coca_RE_cluster -->
<!--   ) %>%  -->
<!--   pivot_longer(cols = -coca_RE_cluster,  -->
<!--                names_to = "cell_type",  -->
<!--                values_to = "cell_count") %>%  -->
<!--   mutate(all = "All") %>%  -->
<!--   filter(!is.na(cell_count)) %>%  -->
<!--   ggplot(aes(x = all, y = cell_count, color = coca_RE_cluster))+ -->
<!--   geom_boxplot()+ -->
<!--   labs(x = "", y = "Cell Count")+ -->
<!--   facet_wrap(cell_type ~ .)+ -->
<!--   theme(axis.text.x = element_blank(), -->
<!--         axis.ticks.x = element_blank()) -->
<!-- ``` -->

### ridgeline
```{r}
library(ggridges)
met_data %>% 
  select(Monocyte : Bmem, Treg,
         CD4, CD8,
         coca_RE_cluster
  ) %>% 
  pivot_longer(cols = -coca_RE_cluster, 
               names_to = "cell_type", 
               values_to = "cell_count") %>% 
  mutate(all = "All") %>% 
  filter(!is.na(cell_count)) %>% 
  ggplot(aes(x=cell_count, y=cell_type, fill=coca_RE_cluster, color=coca_RE_cluster)) +
  geom_density_ridges(alpha=0.3, scale = 0.95, draw_baseline = FALSE)+
  labs(x= "Cell Count", y= "")+
  scale_y_discrete(expand = c(0, 0.2))+
  scale_color_manual(values = c("red", "blue"),
                     aesthetics = c("colour", "fill"))+
  theme(legend.position = "bottom")
```

a little prettier but need to change the cell name yet
```{r}
met_data %>% 
  select(Monocyte : Bmem, Treg,
         CD4, CD8,
         coca_RE_cluster
  ) %>% 
  pivot_longer(cols = -coca_RE_cluster, 
               names_to = "cell_type", 
               values_to = "cell_count") %>% 
  mutate(all = "All") %>% 
  filter(!is.na(cell_count)) %>% 
  ggplot(aes(x=cell_count, y=cell_type, fill=coca_RE_cluster, color=coca_RE_cluster)) +
  geom_density_ridges(alpha=0.2, scale = 0.95, draw_baseline = FALSE)+
  labs(x= "Cell Count", y= "")+
  scale_y_discrete(expand = c(0, 0.2))+
  scale_color_manual(values = c("red", "blue"),
                     aesthetics = c("colour", "fill"),
                     labels = c("Active", "Repressed"), 
                     name = "COCA cluster")+
  xlim(-2, 20)+
  theme(legend.position = "bottom")
```
Questions?   
- Do we want all of the cell types?   
- If yes, I would remove naive B cell and just have 1 curve for B cell... 
The density plot is misleading as no naive B cells are detected in reality.   
- I cropped the x axis and put the legend at the bottom. Not sure we can crop it more as the memory B cell has a peak after 15.   
- What are the units?   








